Testing with Grein et al. data

### 2020-07-09 dev/hmm_grein.Rmd
Work on preprocessing

data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  group_by(patient) %>%
  mutate(breathing_imputed = if_else(breathing == "ND", breathing[which.max(day[breathing %in% c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")])], breathing)) %>%
  ungroup() %>%
  mutate(
    breathing = factor(breathing, levels = breathing_levels, ordered = TRUE),
    breathing_imputed = factor(breathing, levels = breathing_levels, ordered = TRUE)
  )

@@

@@

>>>>

data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  mutate(
    breathing = factor(breathing, levels = breathing_levels, ordered = TRUE),
  )

@@

data_grein %>% ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile()

data_grein %>% group_by(breathing) %>% summarise(pocet = n())
mean(data_grein$breathing == "LoO2", na.rm = TRUE)
```

@@

## No NIPPV, improving/worsening/stable

disease_levels_m5 <- c("AA", "O2", "MV", "ECMO")
breathing_levels_m5 <- c("Discharge", disease_levels_m5, "Death")

m5_observed_state_data <- tibble(id = factor(breathing_levels_m5, levels = breathing_levels_m5), is_noisy = FALSE)

m5_hidden_state_data <- m4_hidden_state_data %>% filter(corresponding_obs != "NIPPV") %>% droplevels()

m5_hidden_state_data <- rbind(
  tibble(base = disease_levels_m5) %>% mutate(corresponding_obs = base) %>% 
    crossing(tibble(state_group = c("_improving", "_stable", "_worsening"))) %>% 
    mutate(id = factor(paste0(base, state_group))) %>%
    select(-base, -state_group),
  tibble(id = c("Death","Discharge")) %>% mutate(corresponding_obs = id)
)

m5_rate_data <- rbind(
  tibble(.from = paste0(disease_levels_m5, "_worsening"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA_improving", .to = "Discharge", rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels_m5[2 : length(disease_levels_m5)], "_improving"), 
             .to = paste0(disease_levels_m5[1 : (length(disease_levels_m5) - 1)],"_improving"), 
             rate_group = "improve_one"),
  tibble(.from = paste0(disease_levels_m5[1 : (length(disease_levels_m5) - 1)],"_worsening"), 
             .to = paste0(disease_levels_m5[2 : length(disease_levels_m5)], "_worsening"), 
             rate_group = "worsen_one"),
  tibble(.from = paste0(disease_levels_m5,"_stable"), 
             .to = paste0(disease_levels_m5, "_improving"), 
             rate_group = "to_improving"),
  tibble(.from = paste0(disease_levels_m5,"_stable"), 
             .to = paste0(disease_levels_m5, "_worsening"), 
             rate_group = "to_worsening"),
  tibble(.from = paste0(disease_levels_m5,"_worsening"), 
             .to = paste0(disease_levels_m5, "_stable"), 
             rate_group = "worsening_to_stable"),
  tibble(.from = paste0(disease_levels_m5,"_improving"), 
             .to = paste0(disease_levels_m5, "_stable"), 
             rate_group = "improving_to_stable")
  ) %>%
  mutate(.rate_id = factor(1:n())) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  recode_group("to_improving") %>%
  recode_group("to_worsening") %>%
  recode_group("improving_to_stable") %>%
  recode_group("worsening_to_stable")

series_data_m5 <- series_data %>% mutate(.observed = factor(case_when(
  .observed == "NIPPV" ~ "MV", 
  .observed %in% c("LoO2","HiO2") ~ "O2",
  TRUE ~ as.character(.observed)), 
  levels = breathing_levels_m5, ordered = TRUE))

m5_initial_states <- series_data_m5 %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")

m5_prior <- brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 5)", "b")

m5_formula <- ~ factor(.rate_id) #m4_formula

m5 <- brmshmmdata(m5_formula, series_data_m5, 
                  m5_rate_data, m5_hidden_state_data, m5_initial_states, prior = m5_prior,
                  observed_state_data = m5_observed_state_data)

m5_fit <- brmhmm(m5, cache_file = paste0(cache_dir,"/grein_m5.rds"), control = list(adapt_delta = 0.9))

summary(m5_fit$brmsfit)
```

```{r}
m5_epred_rect <- posterior_epred_rect(m5_fit)
m5_predicted_rect <- posterior_epred_to_predicted(m5_fit, m5_epred_rect)
m5_predicted <- posterior_rect_to_long(m5_fit, m5_predicted_rect)
```

```{r}
m5_predicted_df <- posterior_long_to_df(m5_fit$data, m5_predicted)

step_size <- 6
for(step in 1:ceiling(m5_fit$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(m5_predicted_df, series_id, m5_fit$data) %>% print()

```{r}
pp_check_last_state(m5_fit, m5_predicted_rect)
pp_check_transitions_direction(m5_fit, m5_predicted_rect)
pp_check_transitions_direction(m5_fit, m5_predicted_rect, states_from = 2:6)
```

```{r}
pp_check_transitions(m5_fit, m5_predicted_rect, states_from = 2:4)
pp_check_transitions(m5_fit, m5_predicted_rect, states_from = 5:6)
